GMU MS SWE Handbook\\{\large A Handbook For the Masters of Software Engineering program at George Mason University

### 2024-10-21 demystify.tex
updates

International students, especially those from less well-known universities and smaller countries, often wonder if they should apply to a CS PhD program in the US. Specifically, they are worried about (i) the \textbf{difficulty of getting admitted} and (ii) the \textbf{cost of graduate study} in the US. Part of the reasons is the lack of information and guidance on the admission process in the US (let's face it, public forums like Facebook communities and Reddit are full of confusion and concerns from students).  This section and the rest of this document aim to address these concerns.

First, applying to a good US university \emph{should not} be any harder than at schools in other countries. It might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic, proposal, or adviser in advance (\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries). 
If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.

Second, PhD students in CS \emph{do not} need to worry about funding, especially at good R1
universities in the US. 
If you are admitted, you will almost certainly
receive \emph{full funding} (\autoref{sec:funding}) to support your study. This includes tuition,
health insurance, and stipend (i.e., you get paid for your study). 
In fact, full funding for CS PhD students is the norm in the US, and I'd go as far as to say that if you are not admitted with full funding, you should be cautious about the program and might want to not accept the offer.
Moreover, depending on the university, you may receive additional benefits such as summer pay, laptops, and traveling to conferences and workshops.

@@

Students from top schools with strong research cultures and experience might already know most of the information given here.
My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.

Another goal is to explain the \emph{``why''} behind the admission process, e.g., why LoRs are important, why you should not draft your own LoR, why you should contact professors, etc.

@@

\paragraph{What next?} Make a decision, accept, reject, or defer the offers  (\autoref{sec:accept-postpone-decline}). Ask to meet with potential advisers (e.g., through Open House or separately) and even their students. Ask about computer equipment and software, office space, and other resources; in many cases these will be provided for free by your adviser or department (\autoref{sec:buying-equipment}).

@@

\paragraph{Types of Faculty} There are typically two types of faculty: tenure-line and teaching.  The main difference is their focus: research, teaching, or both.  The type of faculty can affect their ability to serve as your adviser.

\textbf{Tenure-line faculty}, consisting of tenured and tenure-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring PhD students.  They often have a low teaching load, e.g., 1 course per semester. \emph{Tenured} faculty are professors who have been promoted to permanent positions (informally, very hard to fire them).  \emph{Tenure-track} faculty are professors who are on the track to get tenure.  For most schools, tenure-line is the main type of faculty that can serve as a formal adviser of PhD students. \autoref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.

\textbf{Teaching faculty} (sometimes called instructional faculty or professor of practice) mainly focuses on teaching. They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities, e.g., they do not have to worry much about publishing papers or obtaining grants. As such they typically do not have funding (e.g., RA~\autoref{sec:funding}) to support PhD students.
Depending on the university, they might be able to \emph{co-advise} and mentor PhD students with a tenure-line faculty but typically do not formally advise and graduate PhD students.

Note that at many universities do not allow \emph{non-CS profs}, even in tenure-line ones, to solely advise and graduate PhD students in CS (see \autoref{sec:related-fields}).

@@

\section{How long does it take to evaluate an application?}\label{sec:ievaluate}
\sectioninfo{It takes me about 10 minutes to review an application.}

On average, I spend about 10 minutes reviewing each application (less for clear rejections and more for potential acceptances). While this seems short, it is not that difficult to tell if an application is good or bad.  In fact, this
is twice what other faculty spend on average, e.g., Philip Guo spent \href{https://pg.ucsd.edu/PhD-application-tips.htm}{3--5 minutes} per application.

@@

Finally, I enter my decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA). I also recommend very strong candidates for the University Presidential fellowship. Note that while the system has other decision options, eg., admit without funding, provisional admission (e.g., if they need to take some courses), I rarely use them, simply because we already have too many good students and do not have the resources to support them.

@@

In my personal opinion, highlighting URM or LGBTQ+ identity can \emph{be beneficial}, especially if you can articulate how your diverse experiences contribute to diversity and inclusion in academics.  Many universities have a strong commitment to diversity and inclusion and actively
\emph{recruit students from underepresentative backgrounds} (e.g., some \href{https://cse-climate.engin.umich.edu/reports/climate-dei-reports/cse-climate-and-dei-report-2022-2023/#grad-ethnicity}{stats from UMich CSE} and \href{https://www.Mason.edu/news/2022-09/mason-now-top-10-public-university-diversity-innovation-and-cybersecurity-education-us}{Mason} often promotes itself as one of the most diverse universities in the US). Moreover, many scholarships and fellowships are created specifically for URM and LGBTQ+ students, which you should explore if qualify.

@@

Moreover, promoting diversity aligns with the goal of broadening participation in computing (BPC) that is strongly supported within academia (e.g., see the \href{https://plans.bpcnet.org/GeorgeMasonUniversity_ComputerScience_DepartmentalBPCPlan.pdf}{BPC plan at Mason CS}) and government funding agencies such as NSF (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
Also, sometimes providing this information can make it easier for faculty to apply for certain scholarships or fellowships to support you, e.g., from a private donor that has offered a fellowship only to students from certain backgrounds.

Ultimately, this is a personal choice and you do what you feel comfortable.
If you do not feel comfortable disclosing this, you should not feel pressured to do so and you should also ask your LoR writers to not disclose this information.
Note that you can still talk about your experiences and perspectives without mentioning your identity.
If uncertain, you can seek additional guidance from your close professors and collaborators However, it is important to recognize that diversity is strongly valued and celebrated within CS programs in the US, and that you should feel empowered to share your unique experiences and perspectives.

>>>>

International students, especially those from less well-known universities and smaller countries, often wonder if they should apply to a CS PhD program in the US. Specifically, they are worried about (i) the \textbf{difficulty of getting admitted} and (ii) the \textbf{cost of graduate study} in the US. Part of the reasons is the lack of information and guidance on the admission process in the US (let's face it, social places like Facebook and Reddit are full of confusion and contradictory information).  This document aims to address these concerns.

First, applying to a good US university \emph{should not} be any harder than at schools in other countries. It might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic, proposal, or adviser in advance (\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries). It doesn't even require having a CS background (\autoref{sec:non-stem}).
If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.

Second, PhD students in CS \emph{do not} need to worry about funding, especially at good R1
universities in the US. 
If you are admitted, you will almost certainly
receive \emph{full funding} (\autoref{sec:funding}) to support your study. This includes tuition,
health insurance, and stipend (i.e., you get paid for your study). 
For many students, stipends are not only sufficient for them to live but also for their families (e.g., spouse and children) to live comfortably.

Moreover, depending on the university, you may receive additional benefits such as summer pay, laptops, and traveling to conferences and workshops. 
Full funding for CS PhD students is the norm in the US, and I'd go as far as to say that if you are not admitted with full funding, you might want to not accept the offer.

@@

Students from top schools with strong research cultures and experience might already know most of the information given here.
My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
Another goal of this document is to explain the \emph{``why''} behind the admission process, e.g., why LoRs are important, why you should not draft your own LoR, why you should contact professors, etc.

@@

\paragraph{What's next?} Make a decision, accept, reject, or defer the offers  (\autoref{sec:accept-postpone-decline}). Ask to meet with potential advisers (e.g., through Open House or separately) and even their students. Ask about computer equipment and software, office space, and other resources; in many cases these will be provided for free by your adviser or department (\autoref{sec:buying-equipment}).

@@

\paragraph{Types of Faculty} There are typically two types of faculty at an R1 research university: tenure-line and teaching.  The main difference is their focus: research, teaching, or both. Tenure-line faculty are typically more research-focused and can mentor PhD students, while teaching faculty are more teaching-focused and typically do not advise PhD students. Thus when contacting a prof. for research opportunities (\autoref{sec:contact}), you should focus on tenure-line faculty (and not adjunct or emeritus :p).

\textbf{Tenure-line faculty}, consisting of tenured and tenure-track profs., focuses on research, which includes publishing papers, obtaining grants, and mentoring PhD students.  They often have a low teaching load, e.g., 1 course per semester. \emph{Tenured} faculty are professors who have been promoted to permanent positions (informally, very hard to fire them).  \emph{Tenure-track} faculty are professors who are on the track to get tenure.  For most schools, tenure-line is the main type of faculty that can serve as a formal adviser of PhD students. \autoref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.

\textbf{Teaching faculty}, sometimes called instructional faculty or professor of practice, mainly focuses on teaching. They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities, e.g., they do not have to worry much about publishing papers or obtaining grants. As such they typically do not have funding (e.g., RA~\autoref{sec:funding}) to support PhD students.
Depending on the university, they might be able to \emph{co-advise} and mentor PhD students with a tenure-line faculty, but they typically do not formally advise and graduate PhD students.

Note that at many universities do not allow \emph{non-CS profs}, even tenure-line ones, to solely advise and graduate PhD students in CS (see \autoref{sec:related-fields}).

@@

%\section{"Chance me" questions}\label{sec:chance-me}

%Reddit and other grad admission forums are full of ``Chance me'' and ``roast my profile'' questions. These are questions where students post their profiles and ask others to evaluate their chances of getting admitted to their choice of schools.

%These questions are not useful and should be avoided. First, no one knows! Even adcom members can't tell if you would get admitted to their programs or not. As mentioned throughout this guide, the admission process is complex and involves many factors that are not part of your profile, e.g., whether the faculty have funding or are interested in your background. Second, answers to this questions are often not helpful to you.  Faculty and people knowledgeable about the process will not answer these questions, and those who do are often not knowledgeable and might give you false hope or despair. OK, someone say your GRE is low (which does not even matter for most schools (\autoref{sec:standard-tests}) , are you going to retake it?  Or someone says your GPA is low, are you going to apply to a different school?  No, you should apply to the schools you want to go to and see what happens.  If you don't get in, then you can ask for feedback and improve your profile for the next round.

\section{How long does it take to evaluate an application?}\label{sec:ievaluate}
\sectioninfo{It takes me about 10--15 minutes to review an application.}

On average, I spend about 10--15 minutes reviewing each application (less for clear rejections and more for potential acceptances). While this seems short, it is not that difficult to tell if an application is good or bad.  In fact, this
is twice what other faculty spend on average, e.g., Philip Guo spent \href{https://pg.ucsd.edu/PhD-application-tips.htm}{3--5 minutes} per application.

@@

Finally, I enter my decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA). I also recommend very strong candidates for the University Presidential fellowship. Note that while the system has other decision options, eg., admit without funding, provisional admission (e.g., if they need to take some courses), I rarely use them, simply because we want to fully fund our PhD students.

@@

In my personal opinion, highlighting URM or LGBTQ+ identity can \emph{be beneficial}, especially if you can articulate how your diverse experiences contribute to diversity and inclusion in academics.  Many universities have a strong commitment to diversity and inclusion and actively
\emph{recruit students from underepresentative backgrounds} (e.g., some \href{https://cse-climate.engin.umich.edu/reports/climate-dei-reports/cse-climate-and-dei-report-2022-2023/#grad-ethnicity}{stats from UMich CSE} and \href{https://www.Mason.edu/news/2022-09/mason-now-top-10-public-university-diversity-innovation-and-cybersecurity-education-us}{Mason} often tout 
itself as one of the most diverse universities in the US). Moreover, many scholarships and fellowships are created specifically for URM and LGBTQ+ students, which you should explore if qualify.

@@

Moreover, promoting diversity aligns with the goal of broadening participation in computing (BPC) that is strongly supported within academia (e.g., see the \href{https://plans.bpcnet.org/GeorgeMasonUniversity_ComputerScience_DepartmentalBPCPlan.pdf}{BPC plan at Mason CS}) and government funding agencies such as NSF (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
Providing this information can also make it easier for faculty to apply for certain scholarships or fellowships to support you, e.g., from a private donor that has offered a fellowship only to students from certain backgrounds.

Ultimately, this is a personal choice and you do what you feel comfortable.
If you do not feel comfortable disclosing this, you should not feel pressured to do so.
However, if you feel comfortable but are unsure if it is accepted in higher education in the US, then I hope this section helps you understand that it is not only accepted but also valued and celebrated.

%If uncertain, you can seek additional guidance from your close professors and collaborators.

### 2024-10-21 ennv.tex
updates

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1.5in]{geometry}
\usepackage{soul}
% \usepackage[small,compact]{titlesec} %very powerful
\usepackage[most]{tcolorbox}
% \setsecnumdepth{subsection}
% \setcounter{tocdepth}{3}
\usepackage{enumitem}
\usepackage{epigraph}
\usepackage{cite}
\usepackage{caption}
\captionsetup{font=small}
\usepackage{graphicx}
\usepackage{pdfpages}
\usepackage{hyperref}
\usepackage{wrapfig}
\setlength\intextsep{0pt} % remove extra space above and below in-line float
\usepackage{hyperref}
\hypersetup{
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=blue,
  urlcolor=blue,
}
\usepackage{booktabs}

@@

% for cross references
% \renewcommand{\algorithmcflinename}{L.}
\renewcommand{\algorithmautorefname}{Alg.}
\renewcommand{\figureautorefname}{Fig.}
\renewcommand{\tableautorefname}{Tab.}
\renewcommand{\equationautorefname}{Eq.}
\renewcommand{\chapterautorefname}{\S\@gobble}
\renewcommand{\sectionautorefname}{\S\@gobble}
\renewcommand{\subsectionautorefname}{\S\@gobble}
\renewcommand{\appendixautorefname}{\S\@gobble}

@@

\makeatletter
\makeatother

@@

A \emph{neural network} (\textbf{NN})~\cite{Goodfellow-et-al-2016} consists of an input layer, multiple hidden layers, and an output layer. Each layer has a number of neurons, each connected to neurons from previous layers through a predefined set of weights (derived by training the network with data). A \emph{Deep Neural Network} (\textbf{DNN}) is an NN with at least two hidden layers.

@@

The output of a DNN is obtained by iteratively computing  the  values  of  neurons  in  each  layer.
The value of a neuron in the input layer is the input data. The value of a neuron in the hidden layers is computed by applying an \emph{affine transformation} (\autoref{sec:affine}) to values of neurons in the previous layers, then followed by an \emph{activation function} (\autoref{sec:activation}) such as ReLU and Sigmoid. The value of a neuron in the output layer is computed similarly but may skip the activation function.

@@

\chapter{Verification of Neural Networks}\label{sec:verification}
\paragraph{DNN Verification} Given a DNN \(N\) and a property $\phi$, the \emph{DNN verification problem} asks if $\phi$ is a valid property of $N$.
Typically, $\phi$ is a formula of the form $\phi_{in} \Rightarrow \phi_{out}$, where $\phi_{in}$ is a property over the inputs of $N$ and $\phi_{out}$ is a property over the outputs of $N$.
%This form of properties has been used to encode safety and security requirements of DNNs, e.g., safety specifications to avoid collision in unmanned aircraft~\cite{kochenderfer2012next} and \emph{adversarial robustness}~\cite{katz2017towards} properties desired by all DNNs, in which a small input perturbation does not cause major spikes in the DNN's outputs.
A DNN verifier attempts to find a \emph{counterexample} input to $N$ that satisfies $\phi_{in}$ but violates $\phi_{out}$.  If no such counterexample exists, $\phi$ is a valid property of $N$. Otherwise, $\phi$ is not valid and the counterexample can be used to retrain or debug the DNN~\cite{huang2017safety}.

@@

@@

\chapter{Search Algorithms}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1.5in]{geometry}
\usepackage{amsmath}
\usepackage{soul}
% \usepackage[small,compact]{titlesec} %very powerful
\usepackage[most]{tcolorbox}
% \setsecnumdepth{subsection}
% \setcounter{tocdepth}{3}
\usepackage{enumitem}
\usepackage{epigraph}
\usepackage{cite}
\usepackage{caption}
\captionsetup{font=small}
\usepackage{graphicx}
\usepackage{pdfpages}
\usepackage{hyperref}
\usepackage{wrapfig}
\setlength\intextsep{0pt} % remove extra space above and below in-line float
\usepackage{hyperref}
\hypersetup{
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=blue,
  urlcolor=blue,
}
\usepackage{booktabs}

@@

\makeatletter
% for cross references
% \renewcommand{\algorithmcflinename}{L.}
\renewcommand{\algorithmautorefname}{Alg.}
\renewcommand{\figureautorefname}{Fig.}
\renewcommand{\tableautorefname}{Tab.}
\renewcommand{\equationautorefname}{Eq.}
\renewcommand{\chapterautorefname}{\S\@gobble}
\renewcommand{\sectionautorefname}{\S\@gobble}
\renewcommand{\subsectionautorefname}{\S\@gobble}
\renewcommand{\appendixautorefname}{\S\@gobble}

\makeatother

@@

@@

A \emph{neural network} (\textbf{NN})~\cite{Goodfellow-et-al-2016} consists of an input layer, multiple hidden layers, and an output layer. Each layer has a number of neurons, each connected to neurons in the next layer through a predefined set of weights (derived by training the network with data). A \emph{Deep Neural Network} (\textbf{DNN}) is an NN with two or more hidden layers.

@@

The output of an NN is obtained by iteratively computing  the  values  of  neurons  in  each  layer.
The value of a neuron in the input layer is the input data. The value of a neuron in the hidden layers is computed by applying an \emph{affine transformation} (\autoref{sec:affine}) to values of neurons in the previous layers, then followed by an \emph{activation function} (\autoref{sec:activation}) such as ReLU and Sigmoid. The value of a neuron in the output layer is computed similarly but may skip the activation function.

@@

\chapter{Verification of Neural Networks}\label{sec:verification}

\section{The Neural Network Verification Problem and Its Formalization}
\paragraph{DNN Verification} Given a DNN \(N\) and a property $\phi$, the \emph{DNN verification problem} asks if $\phi$ is a valid property of $N$.
Typically, $\phi$ is a formula of the form $\phi_{in} \Rightarrow \phi_{out}$, where $\phi_{in}$ is a property over the inputs of $N$ and $\phi_{out}$ is a property over the outputs of $N$.
%This form of properties has been used to encode safety and security requirements of DNNs, e.g., safety specifications to avoid collision in unmanned aircraft~\cite{kochenderfer2012next} and \emph{adversarial robustness}~\cite{katz2017towards} properties desired by all DNNs, in which a small input perturbation does not cause major spikes in the DNN's outputs.
A DNN verifier attempts to find a \emph{counterexample} input to $N$ that satisfies $\phi_{in}$ but violates $\phi_{out}$.  If no such counterexample exists, $\phi$ is a valid property of $N$. Otherwise, $\phi$ is not valid and the counterexample can be used to retrain or debug the DNN~\cite{huang2017safety}.

@@

\section{Satisfiability and Activation Pattern Search} As with traditional software, DNN verification is often represented as a satisfiability problem.
We thus need to define a formula to represent the network. Typically this formula is a conjunction of constraints representing the affine transformation and activation function of each neuron in the network.
For a network with $L$ layers, $N$ neurons per layer, and ReLU activations this formula is:
\begin{align*}
\alpha = \bigwedge_{\begin{smallmatrix}i \in [1,L]\\ j \in [1,N]\end{smallmatrix}} v_{i,j} = \max \Big( \sum_{k \in [1,N]} (w_{i-1,j,k} \cdot v_{i-1,j}) + b_{i,j}, 0 \Big)
\end{align*}
With this definition DNN verification can be formulated as checking the satisfiability of:
\begin{equation}\label{eq:prob}
  \alpha \land \phi_{in} \land \neg \phi_{out}
\end{equation}
If \autoref{eq:prob} is unsatisfiable, $\phi$ is a valid property of $\mathcal{N}$. Otherwise, $\phi$ is not valid (and the counterexample of the original problem is any satisfying assignment that makes \autoref{eq:prob} true).

For the widely-used ReLU activation problem, this problem becomes a search for \emph{activation patterns}, i.e., boolean assignments representing activation status of neurons, that lead to satisfaction the formula in \autoref{eq:prob}. Modern DNN verification techniques~\cite{bunel2020branch,wang2021beta,ferrari2022complete,duong2024harnessing,duong2023dpllt,ovalbab,katz2019marabou,bak2021nnenum} all adopt this idea and search for satisfying activation patterns.

@@

\chapter{Common Search Algorithms}

\section{Symbolic Search and Constraint solving}
\section{Branch-and-Bound (BnB)}
